### 2017-10-15 mcmc.tex
problems and punctuation

\begin{problem}\label{prob:rosenbrock}
The Rosenbrock density used as a demonstration case for many samplers (see, for example, \citealt{gw}).
Test your sampler on this density:
\begin{eqnarray}
p(\theta_1,\,\theta_2) &\propto&
    \exp\left(-\frac{100\,(\theta_2-{\theta_1}^2)^2+(1-\theta_1)^2}{20}\right)
\quad.
\end{eqnarray}
Tune the Gaussian proposal distribution in your \MH\ \MCMC\ sampler to sample this density efficiently.
What autocorrelation time do you get?
Compare to what \project{emcee} gets.
\end{problem}

@@

If you suspect, or if it is even possible, that your problem is badly
multi-modal, then you will have to start at multiple points in
parameter space and compare the resulting chains.
By ``badly multi-modal'' we mean that there are peaks in the posterior
pdf that are connected by low-enough valleys that it is very unlikely
or takes a long time for a walker to traverse from one peak to the
next.
If this might be an issue, then the best diagnosis is to start many
chains in parallel and check that they lead to identical (or very
similar) posterior inferences (same pdf location and shape and
statistics).

@@

But it is worth remembering that just as \MCMC\ is not a good optimizer
(generically, samples will not lie close to the maximum of the posterior pdf),
it is also not a good \emph{search algorithm}.
There is no sense in which (standard) \MCMC\ methods are efficient at
searching, or engineered to search, all of parameter space.
If you really need to check all of parameter space, you should put
\MCMC\ aside, and do an exhaustive search.\note{You might want to open a
  bottle of wine or two while the code runs.}

@@

\section{Results, error bars, and figures}\label{sec:results}
\nopagebreak
For a committed probabilistic scientist---frequentist or
Bayesian---there is no ``answer'', there are only probability density
functions (pdfs).  Any time in a paper or in an email or in a conversation
we say what ``the answer is''---even if we say it with an error
bar---we have departed the probabilistic program.  There are many
principled ways to depart the program; in another \documentname, we
discuss the economic explanation of how we can make hard decisions in
the context of probabilistic reasoning\note{We intend to write a subsequent
  \documentname\ in this series about these issues.  The idea is that you
  can make principled decisions by optimizing the expected utility under
  the posterior, given the data.  This is a great idea!  Of course we also
  have very deep, fundamental reasons that you \emph{can't know your utility
    precisely}.  The big issue is that your only sensible utility involves
  an integral out to the ``long term'', and the long term (by definition)
  includes outcomes that are outside your present-day quantitative model.}
But without going into that economic model, it is fair to say that a
substantial problem with being a committed probabilist is that when we
\emph{publish}, we are not permitted (not now, at least) to publish a
\emph{probability distribution over publications}; we have to publish
\emph{one single, deterministic text}; we have to make a decision about what
to write in the title, the abstract, the tables, figures, and results
section.  Given a \MCMC\ sampling of the posterior pdf, what do we
report as our \emph{results}?

>>>>

\begin{problem}\label{prob:rosenbrock}
The Rosenbrock density used as a demonstration case for many samplers (see, for example, \citealt{gw}).
Test your sampler on this density:
\begin{eqnarray}
f(\theta_1,\,\theta_2) &=&
    \exp\left(-\frac{100\,(\theta_2-{\theta_1}^2)^2+(1-\theta_1)^2}{20}\right)
\quad.
\end{eqnarray}
Tune the Gaussian proposal distribution in your \MH\ \MCMC\ sampler to sample this density efficiently.
What autocorrelation time do you get?
Compare to what \project{emcee}\note{Available at \url{http://dfm.io/emcee}.} gets.
\end{problem}

@@

If you suspect, or if it is even possible, that your problem is badly
multi-modal, then you will have to start at multiple points in
parameter space and compare the resulting chains.
By ``badly multi-modal'' we mean that there are peaks in the posterior
pdf that are connected by low-enough valleys that it is very unlikely
or takes a long time for a walker to traverse from one peak to the
next.
If this might be an issue, then the best diagnosis is to start many
chains in parallel with different initializations,
  and check that they lead to identical (or very
similar) posterior inferences (same pdf location and shape and
statistics).

@@

But it is worth remembering that---just as \MCMC\ is not a good optimizer
(generically, samples will not lie close to the maximum of the posterior pdf)---%
it is also not a good \emph{search algorithm}.
There is no sense in which (standard) \MCMC\ methods are efficient at
searching, or engineered to search, all of parameter space.
If you really need to check all of parameter space, you should put
\MCMC\ aside, and do an exhaustive search.\note{You might want to open a
  bottle of wine or two while the code runs.}

@@

\section{Results, error bars, and figures}\label{sec:results}
\nopagebreak
For a committed probabilistic scientist---frequentist or
Bayesian---there is no ``answer'', there are only probability density
functions (pdfs).  In a paper or in an email or in a conversation
we might say what ``the answer is'', but even if we say it with an error
bar, we have departed the probabilistic program.  There are many
principled ways to depart the program; in another forum, we hope to say more
about the economic explanation of how we can make hard decisions in
the context of probabilistic reasoning\note{The idea is that you
  can make principled decisions by optimizing the expected utility under
  the posterior, given the data.  This is a great idea!  Of course we also
  have very deep, fundamental reasons that you \emph{can't know your utility
    precisely}.  The big issue is that your only sensible utility involves
  an integral out to the ``long term'', and the long term (by definition)
  includes outcomes that are outside your present-day quantitative model.}
But without going into that economic model, it is fair to say that a
substantial problem with being a committed probabilist is that when we
\emph{publish}, we are not permitted (not now, at least) to publish a
\emph{probability distribution over publications}; we have to publish
\emph{one single, deterministic text}; we have to make a decision about what
to write in the title, the abstract, the tables, figures, and results
section.  Given a \MCMC\ sampling of the posterior pdf, what do we
report as our \emph{results}?
